Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-05-11 WholePlotTrajectories.Rmd
code bidouille

In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, through the fluxes of light, heat and water, and the biotic interaction and competitive pressure *(Goulamoussene Biogeosciences)*. 
The cornerstone of tropical forests ecology is then to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a]. 
For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density *(e.g. PiponiotElife, RutishauserBiotropica)*. 
These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016]. 
However the response of tree species diversity remains unclear, also it is determinant of ecosystems productivity, stability and functioning [@Tilman2014, Liang2016Science] and would be most probably impacted by post-disturbance environmental changes [@BaralotoJAE].

In the short-term disturbance demonstrated negligible or even positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a]. 
Validations of the IDH in the long term, however, remain scarce and mainly rely on the analysis of the rough species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and diversity profiles, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
Furthermore, the functional approach accounting for differenes species biological attributes would be insightful as it reveals the species role in the post-disturbance trajectory of ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
In that respect major functional traits related to species ecology and performance were largely adopted through an integrative framework [@Diaz2005;@Villeger2008a].
The functional trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
In disturbed tropical rainforests, this framework translates into the fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008] with a shift from "conservative" slow-growing species dealing with scarce resources that dominate before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].

>>>>

In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, through the fluxes of light, heat and water, and the biotic interaction and competitive pressure [@Goulamoussene2017]. 
The cornerstone of tropical forests ecology is then to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a]. 
For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
However the response of tree species diversity remains unclear, also it is determinant of ecosystems productivity, stability and functioning [@Tilman2014, @Liang2016] and would be most probably impacted by post-disturbance environmental changes [@Baraloto2012a].

In the short-term disturbance demonstrated negligible or even positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a]. 
Validations of the IDH in the long term, however, remain scarce and mainly rely on the analysis of the rough species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and diversity profiles, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
Furthermore, the functional approach accounting for differenes species biological attributes would be insightful as it reveals the species role in the post-disturbance trajectory of ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
In that respect major functional traits related to species ecology and performance were largely adopted through an integrative framework [@Diaz2005;@Villeger2008a].
The functional trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
In disturbed tropical rainforests, this framework translates into the fostering of disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
It results in a shift from "conservative" slow-growing species dealing with scarce resources that dominate before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
